Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers

### 2016-06-27 notes/paper/paper.tex
Added timing test figure to the paper.

\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage[english]{babel}
\usepackage{pinyin}
\usepackage{natbib}
\usepackage{notoccite}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage{floatrow}
\usepackage{amsmath}
%
\begin{document}
%
% paper title
\title{Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers}
%
\author{Craig~Euler, Bryan~Juarez, Melissa~Flores, C.T.~Lin}
%
\maketitle
%
\begin{abstract}
Body worn sensors have been used in the market for many years now.
Many of which are used for activity monitoring or fitness.
In this paper, we make use of a matched-filtering algorithm for real-time activity classification with data obtained from three-axis accelerometers worn by the user.
We achieve invariance to sensor orientation with dimensionality reduction.
We make use of an instance-based learning algorithm to train the device to learn the individual's motion patterns and store that information as activity filters for our matched filter.
We improve computational efficiency with dimensionality reduction, decimation in time, and pre-processing of our references post training for use in the matched-filter.
An assessment of our processing time is provided with runtime experiments on the MSP432 microcontroller which shows feasibility of real-time processing capability.
\end{abstract}
%
\section{Introduction}
Various methods for activity classification using body-worn sensors exist.
Custom decision tree, automatically generated decision tree, and artificial neural networks \cite{parkka_ermes_korpipaa_mantyjarvi_peltola_korhonen_2006} have been explored as well as other frequency based classification methods \cite{sharma_purwar_lee_lee_chung_2008}.
In this paper, we make use of a matched-filter method to identify an individual\textquotesingle s activity in real-time.
The matched-filter method with use of thresholding and comparing the correlations of various filters is a simple method to use and have been studied in the past \cite{giannakis_tsatsanis_1990}.
Many of the algorithms and software structure were tailored to computational efficiency to accommodate the resource limitations on the MSP432 microprocessor.
Having an a priori knowledge of the activity the individual is performing is useful for many current fitness algorithms.
Devices such as the Fit-Bit, Apple watch, Moov, Jawbone, etc. usually have the individual actively input the activity they are performing.

@@

The “Analog Discovery 100MSPS USB Oscilloscope and Logic Analyzer” oscilloscope was used for measuring the timing of the algorithms on the microcontroller.
When testing the execution speed, one of the General-Purpose Input/Output (GPIO) pins was connected to a Light-emitted diode (LED) that activates at 3.3 Volts (set high).
This is activated immediately before the beginning of decimation (start of algorithm processing) and deactivated immediately at the end of the matched-fitlering (end of algorithm processing).
The execution time is then recorded on the oscilloscope to determine how long the GPIO was set high.

%
\begin{figure}[!ht]
   \centering
   \subfloat[\label{fig:multiple_comparison}]{\includegraphics[width=0.45\textwidth]{more_time_plots.png}}
   \quad
   \subfloat[\label{fig:timing_refs}]{\includegraphics[width=0.45\textwidth]{one_and_two_references.png}}
   \caption{Timing results with varying sampling frequencies, time-window lengths, and number of references used for the matched-filter on the MSP432 microcontroller processor. The time window used for (b) is 4.0 seconds.}
   \label{fig:timing}
\end{figure}
%
Randomized data was used for the purpose of analysing the throughput performance on the device while varying the parameters.
Figure \ref{fig:multiple_comparison} shows execution time with varying time-window lengths and sampling frequencies while maintaining two references used in the matched-filter process while operating at decimation $2$.
Because the FFT dominates the processing load, and was optimized for even powered factors, one can observe better throughput at radix 2 valued time-window lengths and sampling frequencies followed by other even valued time lengths and sampling frequencies. Both $62$Hz and $50$Hz have only one factor of $2$ hence the expectation that a sampling frequency of $62$Hz processing time is longer than the $50$Hz.
Figure \ref{fig:timing_refs} shows execution time over a $4.0$ second window at decimation $8$ while varying the sampling frequency and number of references. From this, it's observed that matched-filtering accounts for roughly $80$\% of the processing while pre-processing accounts for the other $20$\%.
%
\section{Conclusion}
conclusion text.
%
\appendices
%
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi
%
\bibliographystyle{plain}
\bibliography{reference}
\end{document}

>>>>

\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage[english]{babel}
\usepackage{pinyin}
\usepackage{natbib}
\usepackage{notoccite}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage{floatrow}
\usepackage{amsmath}
%
\begin{document}
%
% paper title
\title{Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers}
%
\author{Craig~Euler, Bryan~Juarez, Melissa~Flores, C.T.~Lin}
%
\maketitle
%
\begin{abstract}
Body worn sensors have been used in the market for many years now.
Many of which are used for activity monitoring or fitness.
In this paper, we make use of a matched-filtering algorithm for activity classification with data obtained from three-axis accelerometers worn by the user.
We show real-time processing capability on the MSP432 microcontroller hardware.
We achieve invariance to sensor orientation with dimensionality reduction.
We make use of an instance-based learning algorithm to train the device to learn the individual's motion patterns and store that information as activity filters for our matched filter.
We improve computational efficiency with dimensionality reduction, decimation in time, and pre-processing of our references post training for use in the matched-filter.
An assessment of our processing time is provided with runtime experiments on the MSP432 microcontroller which shows feasibility of real-time processing capability.
\end{abstract}
%
\section{Introduction}
Various methods for activity classification using body-worn sensors exist.
Custom decision tree, automatically generated decision tree, and artificial neural networks \cite{parkka_ermes_korpipaa_mantyjarvi_peltola_korhonen_2006} have been explored as well as other frequency based classification methods \cite{sharma_purwar_lee_lee_chung_2008}.
In this paper, we make use of a matched-filter method to identify an individual\textquotesingle s activity for the purpose of real-time processing.
The matched-filter method with use of thresholding and comparing the correlations of various filters is a simple method to use and have been studied in the past \cite{giannakis_tsatsanis_1990}.
Many of the algorithms and software structure were tailored to computational efficiency to accommodate the resource limitations on the MSP432 microprocessor.
Having an a priori knowledge of the activity the individual is performing is useful for many current fitness algorithms.

@@

\begin{figure}[!ht]
   \centering
   \includegraphics[width=0.45\textwidth]{32Hz_4sec_8downsample_2Reference_Pulse.PNG}
   \caption{Test setup}
   \label{fig:time_measure}
\end{figure}
%
The “Analog Discovery 100MSPS USB Oscilloscope and Logic Analyzer” oscilloscope was used for measuring the timing of the algorithms on the microcontroller.
When testing the execution speed, one of the General-Purpose Input/Output (GPIO) pins was connected to a Light-emitted diode (LED) that activates at 3.3 Volts (set high).
This is activated immediately before the beginning of decimation (start of algorithm processing) and deactivated immediately at the end of the matched-fitlering (end of algorithm processing).
The execution time is then recorded on the oscilloscope to determine how long the GPIO was set high.

\begin{figure}[!ht]
   \centering
   \subfloat[\label{fig:multiple_comparison}]{\includegraphics[width=0.45\textwidth]{more_time_plots.png}}
   \quad
   \subfloat[\label{fig:timing_refs}]{\includegraphics[width=0.45\textwidth]{one_and_two_references.png}}
   \caption{Timing results with varying sampling frequencies, time-window lengths, and number of references used for the matched-filter on the MSP432 microcontroller processor. The time window used for (b) is 4.0 seconds.}
   \label{fig:timing}
\end{figure}
%
Randomized data was used for the purpose of analysing the throughput performance on the device while varying the parameters.
Figure \ref{fig:multiple_comparison} shows execution time with varying time-window lengths and sampling frequencies while maintaining two references used in the matched-filter process while operating at decimation $2$.
Because the FFT dominates the processing load, and was optimized for even powered factors, one can observe better throughput at radix 2 valued time-window lengths and sampling frequencies followed by other even valued time lengths and sampling frequencies. Both $62$Hz and $50$Hz have only one factor of $2$ hence the expectation that a sampling frequency of $62$Hz processing time is longer than the $50$Hz.
Figure \ref{fig:timing_refs} shows execution time over a $4.0$ second window at decimation $8$ while varying the sampling frequency and number of references. From this, it's observed that matched-filtering accounts for roughly $80$\% of the processing while pre-processing accounts for the other $20$\%.
%
\section{Conclusion}
conclusion text.
%
\appendices
%
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi
%
\bibliographystyle{plain}
\bibliography{reference}
\end{document}
